A general model for the evolution of nuptial gift-giving: Supporting Information

### 2024-12-20 SI.Rmd
Update last comments based on round of revisions and Evol Lett comments

| Information                                                   | Page |
|---------------------------------------------------------------|------|
| S1: Alternative derivation of male fitness threshold          |  2   |
| S2: Operational sex ratio                                     |  4   |
| S3: Sensitivity analysis of parameters in IBM                 |  5   |
| S4: Separate evolution of male search and female choice       |  12  |
| References                                                    |  13  |

@@

S1: Alternative derivation of male fitness threshold
------------------------------------

In the main text, we assumed that males made the decision to search or not search for a nuptial gift.
The expected length of time for which searching males are expected to remain outside of the mating pool is $E[T_{\mathrm{m}}] = \alpha$ (see Model).
Alternatively, we can assume that males search for a period of $T_{\mathrm{m}}$ and spend this full duration of $T_{\mathrm{m}}$ in the time-out phase, even if they succeed in finding a nuptial gift (as we assumed in our IBM).
The probability that a male obtains a nuptial gift during this time is modelled in Eq. 1,

$$P(G) = 1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.$$

$$P(L) = e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.$$

We assume that the fitness increment in reproductive success associated with receiving a nuptial gift versus no nuptial gift are $\lambda(1 + \gamma)$ and $\lambda$, respectively. 
The rate at which males increase their fitness can then be defined as the expected fitness increment from their nuptial gift search divided by $T_{\mathrm{m}}$ plus the time spent in the mating pool waiting to encounter a mate,

$$W_{\mathrm{m}} = \lambda \frac{P(G)\left(1 + \gamma\right) + P(L)}{T_{\mathrm{m}} + \left( \frac{\beta + 1}{R} \right)}.$$

Note that if $R = 0$ or $\lambda = 0$, then, trivially, no change in fitness occurs (since females and males cannot mate or do not produce offspring). 
Fitness is increased by searching for nuptial gifts when $\gamma$ is high, scaled by the expect search time needed to find a nuptial gift. 
A second term on the right-hand side is subtracted, which reflects a loss in fitness proportional to the encounter rate of potential mates in the mating pool.
The conditions under which male inclusive fitness increases by searching for a nuptial gift are found by setting $\partial W_{\mathrm{m}}/\partial T_{\mathrm{m}} = 0$ and solving for $\gamma$ to recover Eq. 2 from the main text,

S2: Operational sex ratio
------------------------------------

S3: Sensitivity analysis of parameters in IBM
------------------------------------

@@

#points(x = alpha_vals, y = g_m, type = "l", lwd = 2, col = "blue");
#points(x = alpha_vals, y = g_f, type = "l", lwd = 2, col = "red", 
#       lty = "dashed");

@@

# points(x = alpha_vals, y = g_m, type = "l", lwd = 2, col = "blue");
# points(x = alpha_vals, y = g_f, type = "l", lwd = 2, col = "red", 
#        lty = "dashed");

@@

@@

References
=============================================

>>>>

| Information                                                   | Page |
|---------------------------------------------------------------|------|
| S1: Sensitivity analysis of parameters in IBM                 |  2   |
| S2: Alternative derivation of male fitness threshold          |  9   |
| S3: Operational sex ratio                                     |  11  |
| S4: Separate evolution of male search and female choice       |  12  |
| S5: Male probability of search trait                          |  13  |
| Reference                                                     |  15  |

@@

S1: Sensitivity analysis of parameters in IBM
------------------------------------

@@

@@

@@

S2: Alternative derivation of male fitness threshold
------------------------------------

In the main text, we assumed that males made the decision to search or not search for a nuptial gift.
The expected length of time for which searching males are expected to remain outside of the mating pool is $E[T_{\mathrm{m}}] = \alpha$ (see Model).
Alternatively, we can assume that males search for a pre-determined period of $T_{\mathrm{m}}$ and spend this full duration of $T_{\mathrm{m}}$ in the time-out phase, even if they succeed in finding a nuptial gift (as we assumed in our IBM).
The probability that a male obtains a nuptial gift during this time is modelled in Eq. 1,

$$G = 1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.$$

$$L = e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.$$

We assume that the fitness associated with receiving a nuptial gift versus no nuptial gift are $\lambda(1 + \gamma)$ and $\lambda$, respectively. 
The rate at which males increase their fitness can then be defined as the expected fitness increase from their nuptial gift search divided by $T_{\mathrm{m}}$ plus the time spent in the mating pool waiting to encounter a mate,

$$W_{\mathrm{m}} = \lambda \frac{G\left(1 + \gamma\right) + L}{T_{\mathrm{m}} + \left( \frac{\beta + 1}{R} \right)}.$$

Note that if $R = 0$ or $\lambda = 0$, then, trivially, no change in fitness occurs (since females and males cannot mate or do not produce offspring). 
Fitness is increased by searching for nuptial gifts when $\gamma$ is high, scaled by the expect search time needed to find a nuptial gift. 
A second term on the right-hand side is subtracted, which reflects a loss in fitness proportional to the encounter rate of potential mates in the mating pool.
The threshold for which male inclusive fitness is not affected by searching for a nuptial gift are found by setting $\partial W_{\mathrm{m}}/\partial T_{\mathrm{m}} = 0$ and solving for $\gamma$ to recover Eq. 2 from the main text,

S3: Operational sex ratio
------------------------------------

@@

S5: Male probability of search trait
-------------------------------------

In the individual-based model of the main text, a focal male $i$ searches for a nuptial gift for a fixed number of $\mathrm{Poisson}(T^{i}_\mathrm{m})$ time steps.
The male will remain outside the mating pool for this number of time steps regardless of whether or not he successfully obtains a nuptial gift.
Here we instead define male search for nuptial gift as an all-or-nothing strategy to align with our analytical model.
Males either search in time-out until they find a nuptial gift, or they do not search at all.
In these simulations, $T^{i}_\mathrm{m}$ is instead defined as the probability that a focal male $i$ searches for a nuptial gift (in the same way that $\rho^{i}$ is defined as the probability that a focal female $i$ rejects a male without a nuptial gift).
For males that decide to search, the time taken until a nuptial gift is obtained is randomly sampled from an exponential distribution with a rate parameter of $\alpha$.
This sampling time is then rounded to the nearest integer to determine the number of time steps that a focal male spends outside the mating pool before obtaining a nuptial gift.

Results are qualitatively identical to those in Figure 3 of the main text (Figure S5.1)

```{r echo = FALSE, eval = TRUE, fig.width = 6, fig.height = 6, fig.cap = "Figure S5.1: Coevolution of male search and female choosiness as a function of nuptial gift search time ($\\alpha$) when male strategy is binary: search or do not search. Points show where the lower 95% confidence interval of male search (blue) and female choosiness (red) exceeds zero, indicating evolution of nuptial gift search or choosiness. Each point includes data from $1000$ replicate simulations with identical starting conditions. Zones are identified that correspond to areas where males search and females are choosy (D), males search but females are not choosy (B), and males do not search and females are not choosy (A), as also depicted in Figures 2 and 3 in the main text. The number of individuals in the population remained at or near carrying capacity of $K = 1000$. In each time step, up to 3000 total pair-wise interactions occurred. Expected female processing time was set to $T_{\\mathrm{f}}=2$ time steps, and $\\gamma$ and $\\alpha$ values in the range [0.0, 2.0] and [0.1, 2.0], respectively, were used."}
SUMM         <- read.csv(file = "inst/Tm_prob.csv");
alpha_vals   <- sort(unique(SUMM[,1]));  
TM_vals      <- tapply(X = SUMM[,3],  INDEX = SUMM[,1], FUN = mean);
m_vals       <- tapply(X = SUMM[,13], INDEX = SUMM[,1], FUN = mean);
beta_vals    <- tapply(X = SUMM[,14], INDEX = SUMM[,1], FUN = mean);
Male_M       <- tapply(X = SUMM[,15], INDEX = SUMM[,1], FUN = mean);
Female_M     <- tapply(X = SUMM[,16], INDEX = SUMM[,1], FUN = mean);
Mf_nupt      <- tapply(X = SUMM[,17], INDEX = SUMM[,1], FUN = mean);
Mf_nonupt    <- tapply(X = SUMM[,18], INDEX = SUMM[,1], FUN = mean);

par(mar = c(5, 5, 1, 1)); 
plot(x = alpha_vals, y = g_f, type = "n", lwd = 2, cex.lab = 1.25, col = "red",
     xlab = expression(paste("Nuptial gift search time (", alpha, ")")),
     ylab = expression(paste("Fitness threshold (", gamma[f], " & ",  gamma[m], 
                             ")")),
     ylim = c(0, 2));

for(i in 1:dim(SUMM)[1]){
  if(SUMM[i, 4] > 0 & SUMM[i, 5] > 0){
    points(x = SUMM[i, 1], y = SUMM[i, 2], cex = 3, pch = 20, col = "blue");
  }
}

for(i in 1:dim(SUMM)[1]){
  if(SUMM[i, 7] > 0 & SUMM[i, 8] > 0){
    points(x = SUMM[i, 1], y = SUMM[i, 2], cex = 2, pch = 20, col = "red");
  }
}
points(x = 0.2, y = 1.9, cex = 14, pch = 22, bg = "red", col = "black");
text(x = 0.2, y = 1.9, labels = "D", cex = 4);
points(x = 1.3, y = 1.5, cex = 14, pch = 22, bg = "blue", col = "black");
text(x = 1.3, y = 1.5, labels = "B", cex = 4, col = "white");
points(x = 1.5, y = 0.5, cex = 14, pch = 22, bg = "white", col = "black");
text(x = 1.5, y = 0.5, labels = "A", cex = 4);
```

Reference
=============================================

### 2024-12-20 SI.tex
Update last comments based on round of revisions and Evol Lett comments

\begin{longtable}[]{@{}ll@{}}
\toprule
Information & Page \\
\midrule
\endhead
S1: Alternative derivation of male fitness threshold & 2 \\
S2: Operational sex ratio & 4 \\
S3: Sensitivity analysis of parameters in IBM & 5 \\
S4: Separate evolution of male search and female choice & 12 \\
References & 13 \\
\bottomrule
\end{longtable}

@@

\hypertarget{s1-alternative-derivation-of-male-fitness-threshold}{%
\subsection{S1: Alternative derivation of male fitness
threshold}\label{s1-alternative-derivation-of-male-fitness-threshold}}

In the main text, we assumed that males made the decision to search or
not search for a nuptial gift. The expected length of time for which
searching males are expected to remain outside of the mating pool is
\(E[T_{\mathrm{m}}] = \alpha\) (see Model). Alternatively, we can assume
that males search for a period of \(T_{\mathrm{m}}\) and spend this full
duration of \(T_{\mathrm{m}}\) in the time-out phase, even if they
succeed in finding a nuptial gift (as we assumed in our IBM). The
probability that a male obtains a nuptial gift during this time is
modelled in Eq. 1,

\[P(G) = 1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.\]

@@

\[P(L) = e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.\]

We assume that the fitness increment in reproductive success associated
with receiving a nuptial gift versus no nuptial gift are
\(\lambda(1 + \gamma)\) and \(\lambda\), respectively. The rate at which
males increase their fitness can then be defined as the expected fitness
increment from their nuptial gift search divided by \(T_{\mathrm{m}}\)
plus the time spent in the mating pool waiting to encounter a mate,

\[W_{\mathrm{m}} = \lambda \frac{P(G)\left(1 + \gamma\right) + P(L)}{T_{\mathrm{m}} + \left( \frac{\beta + 1}{R} \right)}.\]

@@

Note that if \(R = 0\) or \(\lambda = 0\), then, trivially, no change in
fitness occurs (since females and males cannot mate or do not produce
offspring). Fitness is increased by searching for nuptial gifts when
\(\gamma\) is high, scaled by the expect search time needed to find a
nuptial gift. A second term on the right-hand side is subtracted, which
reflects a loss in fitness proportional to the encounter rate of
potential mates in the mating pool. The conditions under which male
inclusive fitness increases by searching for a nuptial gift are found by
setting \(\partial W_{\mathrm{m}}/\partial T_{\mathrm{m}} = 0\) and
solving for \(\gamma\) to recover Eq. 2 from the main text,

@@

\hypertarget{s2-operational-sex-ratio}{%
\subsection{S2: Operational sex ratio}\label{s2-operational-sex-ratio}}

@@

\hypertarget{s3-sensitivity-analysis-of-parameters-in-ibm}{%
\subsection{S3: Sensitivity analysis of parameters in
IBM}\label{s3-sensitivity-analysis-of-parameters-in-ibm}}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Figure S2.1: Evolution of both male search (blue) and female
choice (red) under different combinations of the mortality rates
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\) (mortality in time-in
and out, respectively). The y-axis is the threshold fitness that leads
to evolution of male search (blue) or female choice (red). The results
show noise, but no correlation between the value of the mortality
parameters and the propensity for male search and/or female choice to
evolve. For each of the \(25 \times 20\) combinations of
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\), 3000 replicate
simulations were run.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Figure S2.3 (\(T_{\mathrm{f}} = 10.0\)): The coevolution of
male search and female choosiness as a function of nuptial gift search
time (\(\alpha\)). Points show where the lower 95\% confidence interval
of female choosiness (red) and male search (blue) exceeds zero,
indicating evolution of choosiness or nuptial gift search. Each point
includes data from 3000 replicate simulations with identical starting
conditions. Up to 3000 interactions occur between individuals in each
time step (\(\psi = 3\)), potentially resulting in a mating interaction.
The number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=10.0\) time steps, and \(\gamma\) and \(\alpha\) values
in the range {[}0.1, 2.0{]} and {[}0.0, 2.0{]}, respectively, were
used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{Figure S2.4 (\(\psi = 1\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 1000
interactions occur between individuals in each time step (\(\psi = 1\)),
potentially resulting in a mating interaction. The number of individuals
in the population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 2.0{]}
and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{Figure S2.5 (\(\psi = 2\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 2000
interactions occur between individuals in each time step (\(\psi = 2\)),
potentially resulting in a mating interaction. The number of individuals
in the population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 2.0{]}
and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Figure S2.6 (\(\psi = 4\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 4000
interactions occur between individuals in each time step, potentially
resulting in a mating interaction (\(\psi = 4\)). The number of
individuals in the population remained at or near carrying capacity of
\(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Figure S2.7 (\(\psi = 6\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 6000
interactions occur between individuals in each time step, potentially
resulting in a mating interaction (\(\psi = 6\)). The number of
individuals in the population remained at or near carrying capacity of
\(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

>>>>

\begin{longtable}[]{@{}ll@{}}
\toprule
Information & Page \\
\midrule
\endhead
S1: Sensitivity analysis of parameters in IBM & 2 \\
S2: Alternative derivation of male fitness threshold & 9 \\
S3: Operational sex ratio & 11 \\
S4: Separate evolution of male search and female choice & 12 \\
S5: Male probability of search trait & 13 \\
Reference & 15 \\
\bottomrule
\end{longtable}

@@

\hypertarget{s1-sensitivity-analysis-of-parameters-in-ibm}{%
\subsection{S1: Sensitivity analysis of parameters in
IBM}\label{s1-sensitivity-analysis-of-parameters-in-ibm}}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{Figure S2.1: Evolution of both male search (blue) and female
choice (red) under different combinations of the mortality rates
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\) (mortality in time-in
and out, respectively). The y-axis is the threshold fitness that leads
to evolution of male search (blue) or female choice (red). The results
show noise, but no correlation between the value of the mortality
parameters and the propensity for male search and/or female choice to
evolve. For each of the \(25 \times 20\) combinations of
\(\mu_{\mathrm{in}}\) and \(\mu_{\mathrm{out}}\), 3000 replicate
simulations were run.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{Figure S2.3 (\(T_{\mathrm{f}} = 10.0\)): The coevolution of
male search and female choosiness as a function of nuptial gift search
time (\(\alpha\)). Points show where the lower 95\% confidence interval
of female choosiness (red) and male search (blue) exceeds zero,
indicating evolution of choosiness or nuptial gift search. Each point
includes data from 3000 replicate simulations with identical starting
conditions. Up to 3000 interactions occur between individuals in each
time step (\(\psi = 3\)), potentially resulting in a mating interaction.
The number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=10.0\) time steps, and \(\gamma\) and \(\alpha\) values
in the range {[}0.1, 2.0{]} and {[}0.0, 2.0{]}, respectively, were
used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Figure S2.4 (\(\psi = 1\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 1000
interactions occur between individuals in each time step (\(\psi = 1\)),
potentially resulting in a mating interaction. The number of individuals
in the population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 2.0{]}
and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{Figure S2.5 (\(\psi = 2\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 2000
interactions occur between individuals in each time step (\(\psi = 2\)),
potentially resulting in a mating interaction. The number of individuals
in the population remained at or near carrying capacity of \(K = 1000\).
Expected female processing time was set to \(T_{\mathrm{f}}=2\) time
steps, and \(\gamma\) and \(\alpha\) values in the range {[}0.0, 2.0{]}
and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{Figure S2.6 (\(\psi = 4\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 4000
interactions occur between individuals in each time step, potentially
resulting in a mating interaction (\(\psi = 4\)). The number of
individuals in the population remained at or near carrying capacity of
\(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Figure S2.7 (\(\psi = 6\)): The coevolution of male search and
female choosiness as a function of nuptial gift search time
(\(\alpha\)). Points show where the lower 95\% confidence interval of
where male search (blue) exceeds zero, indicating evolution of
choosiness or nuptial gift search. Each point includes data from 3000
replicate simulations with identical starting conditions. Up to 6000
interactions occur between individuals in each time step, potentially
resulting in a mating interaction (\(\psi = 6\)). The number of
individuals in the population remained at or near carrying capacity of
\(K = 1000\). Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\hypertarget{s2-alternative-derivation-of-male-fitness-threshold}{%
\subsection{S2: Alternative derivation of male fitness
threshold}\label{s2-alternative-derivation-of-male-fitness-threshold}}

In the main text, we assumed that males made the decision to search or
not search for a nuptial gift. The expected length of time for which
searching males are expected to remain outside of the mating pool is
\(E[T_{\mathrm{m}}] = \alpha\) (see Model). Alternatively, we can assume
that males search for a pre-determined period of \(T_{\mathrm{m}}\) and
spend this full duration of \(T_{\mathrm{m}}\) in the time-out phase,
even if they succeed in finding a nuptial gift (as we assumed in our
IBM). The probability that a male obtains a nuptial gift during this
time is modelled in Eq. 1,

\[G = 1 - e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.\]

@@

\[L = e^{-\frac{1}{\alpha}T_{\mathrm{m}}}.\]

We assume that the fitness associated with receiving a nuptial gift
versus no nuptial gift are \(\lambda(1 + \gamma)\) and \(\lambda\),
respectively. The rate at which males increase their fitness can then be
defined as the expected fitness increase from their nuptial gift search
divided by \(T_{\mathrm{m}}\) plus the time spent in the mating pool
waiting to encounter a mate,

\[W_{\mathrm{m}} = \lambda \frac{G\left(1 + \gamma\right) + L}{T_{\mathrm{m}} + \left( \frac{\beta + 1}{R} \right)}.\]

@@

Note that if \(R = 0\) or \(\lambda = 0\), then, trivially, no change in
fitness occurs (since females and males cannot mate or do not produce
offspring). Fitness is increased by searching for nuptial gifts when
\(\gamma\) is high, scaled by the expect search time needed to find a
nuptial gift. A second term on the right-hand side is subtracted, which
reflects a loss in fitness proportional to the encounter rate of
potential mates in the mating pool. The threshold for which male
inclusive fitness is not affected by searching for a nuptial gift are
found by setting \(\partial W_{\mathrm{m}}/\partial T_{\mathrm{m}} = 0\)
and solving for \(\gamma\) to recover Eq. 2 from the main text,

@@

\hypertarget{s3-operational-sex-ratio}{%
\subsection{S3: Operational sex ratio}\label{s3-operational-sex-ratio}}

@@

@@

\hypertarget{s5-male-probability-of-search-trait}{%
\subsection{S5: Male probability of search
trait}\label{s5-male-probability-of-search-trait}}

In the individual-based model of the main text, a focal male \(i\)
searches for a nuptial gift for a fixed number of
\(\mathrm{Poisson}(T^{i}_\mathrm{m})\) time steps. The male will remain
outside the mating pool for this number of time steps regardless of
whether or not he successfully obtains a nuptial gift. Here we instead
define male search for nuptial gift as an all-or-nothing strategy to
align with our analytical model. Males either search in time-out until
they find a nuptial gift, or they do not search at all. In these
simulations, \(T^{i}_\mathrm{m}\) is instead defined as the probability
that a focal male \(i\) searches for a nuptial gift (in the same way
that \(\rho^{i}\) is defined as the probability that a focal female
\(i\) rejects a male without a nuptial gift). For males that decide to
search, the time taken until a nuptial gift is obtained is randomly
sampled from an exponential distribution with a rate parameter of
\(\alpha\). This sampling time is then rounded to the nearest integer to
determine the number of time steps that a focal male spends outside the
mating pool before obtaining a nuptial gift.

Results are qualitatively identical to those in Figure 3 of the main
text (Figure S5.1)

\begin{figure}
\centering
\includegraphics{SI_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{Figure S5.1: Coevolution of male search and female choosiness
as a function of nuptial gift search time (\(\alpha\)) when male
strategy is binary: search or do not search. Points show where the lower
95\% confidence interval of male search (blue) and female choosiness
(red) exceeds zero, indicating evolution of nuptial gift search or
choosiness. Each point includes data from \(1000\) replicate simulations
with identical starting conditions. Zones are identified that correspond
to areas where males search and females are choosy (D), males search but
females are not choosy (B), and males do not search and females are not
choosy (A), as also depicted in Figures 2 and 3 in the main text. The
number of individuals in the population remained at or near carrying
capacity of \(K = 1000\). In each time step, up to 3000 total pair-wise
interactions occurred. Expected female processing time was set to
\(T_{\mathrm{f}}=2\) time steps, and \(\gamma\) and \(\alpha\) values in
the range {[}0.0, 2.0{]} and {[}0.1, 2.0{]}, respectively, were used.}
\end{figure}

\hypertarget{reference}{%
\section*{Reference}\label{reference}}
\addcontentsline{toc}{section}{Reference}
